---
title: Molecular detection of plant viruses infecting pepper in Oklahoma and their
  subsequent characterization
author: "Caleb Paslay"
date: "`r Sys.Date()`"
output:
  ioslides_presentation: default
  powerpoint_presentation: default
college: University of Tulsa
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
```

## Background 

- How important is pepper-related research?
- Pepper production, health benefits, antimicrobial, economic impact
- Pepper crop -> virus impact -> virus impact on pepper crop

##

```{r}
# Trialing some data available on R
sw <- starwars %>% 
  select("Mass" = mass, "Height" = height, "Name" = name) %>% 
  arrange(desc(Mass)) %>% 
  head(n = 10) #use (n) to change the number of observations 

knitr::kable(sw) #converts the data.frame into a very nice table (use knitr::)
```


## Primary Questions

- What viruses are presently infecting pepper (*Capsicum*) crops throughout Oklahoma? 

- Which pepper cultivars may possess natural resistance or tolerance to viral pathogens?

- How can we prepare growers for the current "climate" of viral presence?


## General Objectives
- field sample screening
- Resistance screening

## Objectives pertaining to field samples
- Collect field samples from various locations throughout OK
- Screen for Pepper mild mottle virus (PMMoV)
- Use NGS to detect unreported or novel viruses
- Complete genome assembly of viruses obtained
- Characterize viruses identified using NGS by RT-PCR
- Sequence confirmation
- Phylogenetic analyses 

## Objectives pertaining to resistance screening
- Screen for PMMoV against various cultivars

## Preliminary Results

```{r}
#field survey data (samples collected, screened, etc.)
```

```{r}
#field screening data (virus incidence and detection)
```

```{r}
#NGS sequencing data (ilumina sequences obtained)
```

## Preliminary Results
### **Currently identified viruses**

1. Alfalfa mosaic virus (AMV)
2. Beet Curly top virus (BCTV)
3. Beet western yellows virus (BWYV)
4. Cucumber mosaic virus (CMV)
5. Pepper yellow dwarf virus (PepYDV)
6. Pepper mild mottle virus (PMMoV)
7. Potato yellow dwarf virus (PYDV)
8. Tomato yellow leaf curl virus (TYLCV)

## Preliminary Results
### **RNA viruses (Riboviria)**
1. Alfalfa mosaic virus (AMV)
2. Beet western yellows virus (BWYV)
3. Cucumber mosaic virus (CMV)
4. Pepper mild mottle virus (PMMoV)
5. Potato yellow dwarf virus (PYDV)

## Preliminary Results
### **DNA viruses (Dnaviria)**
1. Beet Curly top virus (BCTV)
2. Pepper yellow dwarf virus (PepYDV)
3. Tomato yellow leaf curl virus (TYLCV)





## Smaller investigations
- Inoculations 
- Seed transmission

