---
title: "Introduction to R: Part 1"
author: "Caleb Paslay"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty: 
    theme: hpstr
---


[R introduction](https://www.w3schools.com/r/r_intro.asp)


## What is R? 

R and R studio are programming devices (languages) used to visualize and represent data. They are also used for statistical analyses and a breadth of many other things. 

## Why use R? 

1. It is a great resource for data analysis, data visualization, data science and machine learning

2. It provides many statistical techniques (such as statistical tests, classification, clustering and data reduction) 

3. It is easy to draw graphs in R, like pie charts, histograms, box plot, scatter plot, etc++ 

4. It works on different platforms (Windows, Mac, Linux) 

5. It is open-source and free 

6. It has a large community support 

7. It has many packages (libraries of functions) that can be used to solve different problems 

## Installation of R
First we will need to install R. This can be done using various computers (mac, windows, linux).

```{r}
# Insert R code (Command + option + I)
```

We can use the R studio console like a basic calculator. 
```{r message = TRUE}
5 + 5
```

If we would like to output a word, we need to use "". 
```{r}
"Hello World"
```

To output numbers we simply input the numbers. 
```{r}
1
```

Often, other coding languages will use the print() function to output information. 
```{r}
print("hello world")
```

The print function is necessary in many cases, so its best to use it consistently. Here is an example of it being used for code that we will learn later:

```{r}
for (y in 2000:2010) {
  print(y)
}
```

We can use # to comment on our code.
```{r}
#using a for...in loop 
for (y in 2000:2010) {
  print(y) #this is how we print the output to the console
}
```

A powerful aspect of R studio is the ability to create and manipulate variables. 

Name and age are variables and "John" and 23 are values.
```{r}
name <- "John" #assigning a character variable
age <- 23 
name #output "john"
age #output 23
print(name)
```

We can concatenate elements or join elements. 
```{r}
text <- "Jared"
paste(text, "is awesome!")
```

We can begin to build multiple elements and variables. 
```{r}
text1 <- "R"
text2 <- "is awesome!"
paste(text1, text2)
```

When working with numbers, we use the "+" operator. 
```{r}
num1 <- 10
num2 <- 5
num1 + num2
```

We can assign the same value to multiple variables in one line of code.
```{r}
var1 <- var2 <- var3 <- c("orange")
print(var1)
```

## Legal variable names:
myvar <- "John"
my_var <- "John"
myVar <- "John"
MYVAR <- "John"
myvar2 <- "John"
.myvar <- "John"

## Illegal variable names:
2myvar <- "John"
my-var <- "John"
my var <- "John"
_my_var <- "John"
my_v@ar <- "John"
TRUE <- "John"

Understand the concepts of character types. 

1. numeric - (10.5, 55, 787)

2. integer - (1L, 55L, 100L, where the letter "L" declares this as an integer)

3. complex - (9 + 3i, where "i" is the imaginary part)

4. character (a.k.a. string) - ("k", "R is exciting", "FALSE", "11.5")

5. logical (a.k.a. boolean) - (TRUE or FALSE)

```{r}
x <- 10.5 #numeric type
y <- 100L #integer
z <- 9i + 3 # complex
xx <- "R is exciting" #character string
yy <- TRUE #logical/boolean

class(x) #tells us the type
class(y)
class(z)
class(xx)
class(yy)
```

## The different types of numbers

1. Numeric (most commonly used)
```{r}
x <- 10.5 
y <- 11
x
y
```
2. Integer (used when you are certain that decimals will not be used)
```{r}
x <- 50L #must use the letter L to specify 
y <- 100L
x
y
```
3. Complex
```{r}
x <- 3i + 5
y <- 5i
x
y
```

We can convert from one type to another used base R code functions.
```{r}
x <- 1L # integer
y <- 2 # numeric
# convert from integer to numeric:
a <- as.numeric(x)
# convert from numeric to integer:
b <- as.integer(y)
x
y
class(a)
class(b)
```

We briefly introduced the use of R as a basic calculator, but R also offers math functions.
```{r}
max(c(10, 5, 12)) #returns the maximum within the group.
min(c(10,5,12)) #returns the minimum within the group.
sqrt(16) #returns the squareroot.
abs(-4.8) #returns the absolute (positive) value.
ceiling(4.4) #rounds the value up to the nearest integer.
floor(4.4) #rounds the value down to the nearest integer. 
```

## Working with strings of characters


```{r}
str <- "Hello world"
# (nchar) number of characters in a string. 
nchar(str) 

# (grepl) used to check if a character is present in a string.
grepl("H", str) 
grepl("W", str)
grepl("X", str)

# (paste) used merge/concatenate two strings
str1 <- "hello"
str2 <- "world"
paste(str1, str2)
```

## Using the escape characters 

```{r}
# Using a \ followed by the character that is needing to be excluded.
str <- "We are the so-called \"Vikings\", from the north."
str
cat(str)
```

\\ backslash

\n new line 

\r carriage return

\t tab

\b backspace


## Boolean Operators (<, >, ==, if, then, or, else)

```{r}
#Boolean is often used to determine if something is TRUE or FALSE
10 < 11 #TRUE
10 > 11 #FALSE
10 == 9 #FALSE
```

We can also compare different variables. 
```{r}
a <- 10 
b <- 15
a > b 
```
We can also use the IF operator 
```{r}
if (b > a){
  print("b is greater than a")
}else{
  print("b is not greater than a")
}
```

## R operators 

### Arethmetic Operators 
```{r}
5 + 5 #addition operator
5 - 5 #subtraction operator 
5 * 5 #multiplication operator
5 / 5 #division operator 
5^2 #exponent operator
5 %% 5 #Modulus (remainder from division)
5 %/% 5 #integer division
```

### Assignment operators 
```{r}
my_var <- 3
my_var <<- 3 #global assigner
3 -> my_varr
```

### R Comparision operators
```{r}
x <- 40
y <- 50

x == y #Equal
x != y #Not equal
x > y #greater than
x < y #less than
x >= y #greater than or equal to
x <= y #less than or equal to
```

### R Logical operators (used to combine conditional statements)
```{r}
#& Element-wise Logical AND operator. It returns TRUE if both elements are TRUE
#&&	Logical AND operator - Returns TRUE if both statements are TRUE
#| Element-wise Logical OR operator. It returns TRUE if one of the statement is TRUE
#|| Logical OR operator. It returns TRUE if one of the statement is TRUE.
#! Logical NOT - returns FALSE if statement is TRUE
```

### R miscellaneous operators
```{r}
x <- 1:10 #Creates a series of numbers in a sequence	
x %in% y #Find out if an element belongs to a vector
#x <- Matrix1 %*% Matrix2 Matrix Multiplication
```






