---
title: "Introduction to R: Part 3"
author: "Caleb Paslay"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty: 
    theme: hpstr
---

### R Vectors

A vector is simply a list of items that are of the same type. To combine the list of items to a vector, use the c() function and separate the items by a comma. In the example below, we create a vector variable called fruits, that combine strings:

Character values

```{r}
fruits <- c("apple", "orange", "grape")
fruits
```

Numerical values

```{r}
numbers <- c(10,20,30)
numbers
```

Numerical values with decimals (sequence of numbers)

```{r}
numbers1 <- 1.5:5.5
numbers1

numbers2 <- 1.5:5.3 #last element is not used
numbers2
```

Vector of logical values

```{r}
log_value <- c(TRUE, FALSE, TRUE)
log_value
length(log_value) #gives us the length of the vector
```

Sorting values in a vector

```{r}
fruit <- c("apple", "berries", "orange", "grape", "pear")
number <- c(1,2,3,4,5,6,7,8,9,10)

sort(fruit) #sorts alphabetically
sort(number) #sorts numerically
```

### Access vectors by using the [] brackets

```{r}
student <- c("aaron", "mike", "haley", "ryan")

#access the first item in a vector
student[2]
```

Access multiple elements

```{r}
student <- c("aaron", "mike", "haley", "ryan")

student[c(1,2)]
```

Access elements by negation

```{r}
student <- c("aaron", "mike", "haley", "ryan")
student[c(-1,-2)] #Not 1
```

Changing an element within the vector

```{r}
student <- c("aaron", "mike", "haley", "ryan")
student[1] <- "rodger" #changing the element

student
```

### Repeating Vectors

```{r}
repeat_each <- rep(c(1,2,3), each =3) #repeat each value
repeat_each 
```

```{r}
repeat_times <- rep(c(1,2,3), times = 3) #repeat sequence of the vector

repeat_times
```

```{r}
repeat_indepent <- rep(c(1,2,3), times = c(5,2,1)) #repeat each value independently

repeat_indepent
```

### Creating sequenced vectors (: or seq)

```{r}
numbers <- 1:20
numbers
```

```{r}
#from: where sequence begins
#to: where sequence stops
#by: interval of the sequence
numbers <- seq(from = 0, to = 20, by = 0.5)
numbers
```


