---
title: "Introduction to R: Part 2"
author: "Caleb Paslay"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty: 
    theme: hpstr
---

## Working with R if...else statements
```{r}
a <- 33
b <- 40
if (b >= a){
  print("b is greater than a!")
}
```

The if..else statement means "If the previous conditions were not true, then try this condition". 

```{r}
a <- 100
b <- 200
if (a > b){
  print("a is less than b!")
}else if (a < b){
    print("a is less than b")
}
```

```{r}
a <- 200
b <- 200
if (b > a){
  print("b is greater than a")
}else if(a == b){
  print("a and b are equal")
}else{ #we can use else without the else if function
  print("a is greater than b")
}
```

```{r}
a <- 200
b <- 300

if (a > b){
  print("a is greater than b")
}else {
print("b is greater than a")
}
```

```{r}
x <- 40
if (x > 10) {
  print("Above ten")
  if (x > 20) {
    print("and also above 20!")
  } else {
    print("but not above 20.")
  }
} else {
  print("below 10.")
}
```

## AND
```{r}
a <- 200
b <- 33
c <- 500

if (a > b & c > b){
  print("both conditions are true")
}
```

## OR
```{r}
a <- 200
b <- 33
c <- 500

if (a > b  | c > b){
  print("b is a lower than 200")
}
```

## Loops 
Loops can execute a block of code as long as a specified condition is reached. Loops are useful because they save time, reduce errors, and they make code more readable.

### R while loops 
With the while loop we can execute a set of statements as long as a condition is TRUE. 
```{r}
i <- 1
while(i > 6) {
  print(i)
  i <- i + 1 #perform as an increment or the loop will continue forever.
}
```

### Breaking the loop
```{r}
i <- 1
while (i < 6) {
  print(i)
  i <- i + 1
  if (i == 4) {
    break #at 4, the loop will be stopped
  }
}
```

### Is the Next iteration
```{r}
i <- 0
while (i < 6){
  i <- i + 1
  if(i == 3){
    next
  }
  print(i)
}
```

### If...else and while loops 
```{r}
dice <- 0
while (dice <= 6) {
  if (dice < 6) {
    print("No Yahtzee")
  } else {
    print("Yahtzee!")
  }
  dice <- dice + 1
}
```

### R for loops (used for iterating over a sequence)
```{r}
for(x in 1:10){
  print(x)
}
```

We can print every item within a list.
```{r}
fruits <- c("apple", "pear", "grape")

for(x in fruits){
  print(x)
}
```

printing the numbers of a dice.
```{r}
dice <- c(1, 2, 3, 4, 5, 6)

for (x in dice) {
  print(x)
}
```

using the break function to stop the loop.
```{r}
fruits <- list("apple", "banana", "cherry")

for (x in fruits) {
  if (x == "cherry") {
    break
  }
  print(x)
}
```

using the next function.
```{r}
fruits <- list("apple", "banana", "cherry")

for (x in fruits) {
  if (x == "banana") {
    next
  }
  print(x)
}
```

```{r}
dice <- 1:6

for(x in dice) {
  if (x == 6) {
    print(paste("The dice number is", x, "Yahtzee!"))
  } else {
    print(paste("The dice number is", x, "Not Yahtzee"))
  }
}
```

Nesting loops.
```{r}
adj <- list("red", "big", "tasty")

fruits <- list("apple", "banana", "cherry")
  for (x in adj) {
    for (y in fruits) {
      print(paste(x, y))
  }
}
```

### Functions 
A function is a block of code which only runs when it is called.
You can pass data, known as parameters, into a function.
A function can return data as a result.

```{r}
my_function <- function(fname) {
  paste(fname, "Green")
}

my_function("Mr")
my_function("Ms")
my_function("Mrs")
```

A functions called arguments must equal the number of arguments in the function. 

```{r}
my_function <- function(fname, lname) {
  paste(fname, lname)
}

my_function("Caleb", "Paslay")
```

Setting default parameters. 
```{r}
my_function <- function(country = "Norway") {
  paste("I am from", country)
}

my_function("Sweden")
my_function("India")
my_function() # will get the default value, which is Norway
my_function("USA")
```

Return values. 
```{r}
my_function <- function(L,W){
  return(L * W) #function for the area of a square or rectangle
}

my_function(2, 4) #plug in the length and width
```

```{r}
my_function <- function(r){
  return(3.14 * r^2)
}

my_function(20)
```

### Nesting functions
There are two ways to create a nested function:
Call a function within another function.
Write a function within a function.
```{r}
nested_function <- function(x,y){
  a <- x + y 
  return(a)
}
nested_function(nested_function(2,2), nested_function(3,3))
```

```{r}
Outer_func <- function(x) {
  Inner_func <- function(y) {
    a <- x + y
    return(a)
  }
  return (Inner_func)
}
output <- Outer_func(3) # To call the Outer_func
output(5)
```

### R global variables
Variables that are created outside of a function are known as global variables. Global variables can be used by everyone, both inside of functions and outside.

```{r}
txt <- "awesome"
my_function <- function() {
  paste("R is", txt) #join or merge
}

my_function()
```

```{r}
txt <- "global variable"
my_function <- function() {
  txt = "fantastic"
  paste("R is", txt)
}

my_function()

txt # print txt
```

```{r}
my_function <- function() {
txt <<- "fantastic" #global variable inside a function
  paste("R is", txt)
}

my_function()

print(txt)
```







